---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
options(scipen=999)
```

## Loading and preprocessing the data

```{r}

library(tidyverse)
df_data <- read_csv(file = "activity/activity.csv")
```


## What is mean total number of steps taken per day?

```{r}
# For this part of the assignment, you can ignore the missing values in the dataset.
# 1. Make a histogram of the total number of steps taken each day
# 2. Calculate and report the mean and median total number of steps taken per day

df_total_steps_per_day <- df_data %>% 
  group_by(date) %>% 
  summarise(total_steps_per_day = sum(steps, na.rm = TRUE))

hist(df_total_steps_per_day$total_steps_per_day, main = "Total number of steps taken each day", xlab = "Total number of steps")
```

>The mean total number of steps taken per day is : **`r round(mean(df_total_steps_per_day$total_steps_per_day), 0)`**  

>The median total number of steps taken per day is : **`r median(df_total_steps_per_day$total_steps_per_day)`**


## What is the average daily activity pattern?

```{r echo=FALSE}
# 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis)
# and the average number of steps taken, averaged across all days (y-axis)
# 2. Which 5-minute interval, on average across all the days in the dataset,
# contains the maximum number of steps?

```

```{r}
df_average_steps_per_day_interval <-  df_data %>% 
  group_by(interval) %>%
  summarise(average_steps_per_day_interval = mean(steps, na.rm = TRUE))

# create the plot
plot(df_average_steps_per_day_interval$interval, df_average_steps_per_day_interval$average_steps_per_day_interval, 
     type = "l",  main = "Average across all the days", xlab = "Interval", ylab = "Average steps",
     col = "red", lwd =1)
```


```{r}
max_steps <- max(df_average_steps_per_day_interval$average_steps_per_day_interval)
  
interval_with_max <- df_average_steps_per_day_interval %>% 
  filter(average_steps_per_day_interval == max_steps) %>% 
  pull(interval)
```

The 5 minute interval with maximum number of steps is: **`r interval_with_max`**


## Imputing missing values

```{r echo=FALSE}
# Note that there are a number of days/intervals where there are missing values
# (coded as NA). The presence of missing days may introduce bias into some
# calculations or summaries of the data.
# 1. Calculate and report the total number of missing values in the dataset
# (i.e. the total number of rows with NAs)
# 2. Devise a strategy for filling in all of the missing values in the dataset. The
# strategy does not need to be sophisticated. For example, you could use
# the mean/median for that day, or the mean for that 5-minute interval, etc.
# 3. Create a new dataset that is equal to the original dataset but with the
# missing data filled in.
# 4. Make a histogram of the total number of steps taken each day and Calculate
# and report the mean and median total number of steps taken per day. Do
# these values differ from the estimates from the first part of the assignment?
# What is the impact of imputing missing data on the estimates of the total
# daily number of steps?
  
```

1. Calculate and report the total number of missing values in the dataset

Total number of rows with NAs: **`r nrow(df_data %>% filter(is.na(steps)))`**

2. Devise a strategy for filling in all of the missing values in the dataset. The
strategy does not need to be sophisticated. For example, you could use
the mean/median for that day, or the mean for that 5-minute interval, etc.

> Calculated the median for each interval to be used to update the dataset

```{r}
df_median_steps_per_interval <- df_data %>% 
  group_by(interval) %>% 
  summarise(interval_median = median(steps, na.rm = TRUE))
```

3. Create a new dataset that is equal to the original dataset but with the
 missing data filled in.

```{r}
df_filled_NAs <- df_data %>% 
  mutate(steps = ifelse(is.na(steps), recode(interval, !!!setNames(df_median_steps_per_interval$interval_median, df_median_steps_per_interval$interval)), steps))

df_filled_NAs %>% DT::datatable()

```

4. Make a histogram of the total number of steps taken each day and Calculate
and report the mean and median total number of steps taken per day. Do
these values differ from the estimates from the first part of the assignment?
What is the impact of imputing missing data on the estimates of the total
daily number of steps?

```{r}

df_updated_steps_per_day <- df_filled_NAs %>% 
  group_by(date) %>% 
  summarise(total_steps_per_day = sum(steps, na.rm = TRUE))

hist(df_updated_steps_per_day$total_steps_per_day, main = "Total number of steps taken each day", xlab = "Total number of steps")
```

>The mean total number of steps taken per day is : **`r round(mean(df_updated_steps_per_day$total_steps_per_day), 0)`**  

>The median total number of steps taken per day is : **`r median(df_updated_steps_per_day$total_steps_per_day)`**

> The mean total number of days increased after imputing missing values by: **150** 


## Are there differences in activity patterns between weekdays and weekends?

```{r echo=FALSE}
# For this part the weekdays() function may be of some help here. Use the dataset
# with the filled-in missing values for this part.
# 1. Create a new factor variable in the dataset with two levels – “weekday”
# and “weekend” indicating whether a given date is a weekday or weekend day.
# 2. Make a panel plot containing a time series plot (i.e. type = "l") of the
# 5-minute interval (x-axis) and the average number of steps taken, averaged
# across all weekday days or weekend days (y-axis). The plot should look
# something like the following, which was creating using simulated data:

```

1. Create a new factor variable in the dataset with two levels – “weekday”
and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
df_data_with_weekday <- df_filled_NAs %>% 
  mutate(weekd = weekdays(as.Date(date)),
         weekday = ifelse(weekd %in% c("Saturday", "Sunday"), "weekend", "weekday"),
         weekday = factor(weekday))
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the
5-minute interval (x-axis) and the average number of steps taken, averaged
across all weekday days or weekend days (y-axis)

```{r}
df_average_no_steps <- df_data_with_weekday %>% 
  group_by(weekd, weekday, interval) %>% 
  summarise(avg_steps = mean(steps))

lattice::xyplot(avg_steps~interval|as.factor(weekday) , data = df_average_no_steps, layout =c(1,2), type = "l", ylab = "Number of steps")

```

