---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
options(scipen=999)
```

## Loading and preprocessing the data

```{r}

library(tidyverse)
df_data <- read_csv(file = "activity/activity.csv")
```


## What is mean total number of steps taken per day?

```{r}
# For this part of the assignment, you can ignore the missing values in the dataset.
# 1. Make a histogram of the total number of steps taken each day
# 2. Calculate and report the mean and median total number of steps taken per day

df_total_steps_per_day <- df_data %>% 
  group_by(date) %>% 
  summarise(total_steps_per_day = sum(steps, na.rm = TRUE))

hist(df_total_steps_per_day$total_steps_per_day, main = "Total number of steps taken each day", xlab = "Total number of steps")
```

>The mean total number of steps taken per day is : **`r round(mean(df_total_steps_per_day$total_steps_per_day), 0)`**  

>The median total number of steps taken per day is : **`r median(df_total_steps_per_day$total_steps_per_day)`**


## What is the average daily activity pattern?

```{r}

# 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis)
# and the average number of steps taken, averaged across all days (y-axis)
# 2. Which 5-minute interval, on average across all the days in the dataset,
# contains the maximum number of steps?

df_average_steps_per_day_interval <-  df_data %>% 
  group_by(interval) %>%
  summarise(average_steps_per_day_interval = mean(steps, na.rm = TRUE))

# create the plot
plot(df_average_steps_per_day_interval$interval, df_average_steps_per_day_interval$average_steps_per_day_interval, 
     type = "l",  main = "Average across all the days", xlab = "Interval", ylab = "Average steps",
     col = "red", lwd =1)
```


```{r}
max_steps <- max(df_average_steps_per_day_interval$average_steps_per_day_interval)
  
interval_with_max <- df_average_steps_per_day_interval %>% 
  filter(average_steps_per_day_interval == max_steps) %>% 
  pull(interval)
```

The 5 minute interval with maximum number of steps is: **`r interval_with_max`**


## Imputing missing values

```{r}
# Note that there are a number of days/intervals where there are missing values
# (coded as NA). The presence of missing days may introduce bias into some
# calculations or summaries of the data.
# 1. Calculate and report the total number of missing values in the dataset
# (i.e. the total number of rows with NAs)
# 2. Devise a strategy for filling in all of the missing values in the dataset. The
# strategy does not need to be sophisticated. For example, you could use
# the mean/median for that day, or the mean for that 5-minute interval, etc.
# 3. Create a new dataset that is equal to the original dataset but with the
# missing data filled in.
# 4. Make a histogram of the total number of steps taken each day and Calculate
# and report the mean and median total number of steps taken per day. Do
# these values differ from the estimates from the first part of the assignment?
# What is the impact of imputing missing data on the estimates of the total
# daily number of steps?
  

```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
# For this part the weekdays() function may be of some help here. Use the dataset
# with the filled-in missing values for this part.
# 1. Create a new factor variable in the dataset with two levels – “weekday”
# and “weekend” indicating whether a given date is a weekday or weekend day.
# 2. Make a panel plot containing a time series plot (i.e. type = "l") of the
# 5-minute interval (x-axis) and the average number of steps taken, averaged
# across all weekday days or weekend days (y-axis). The plot should look
# something like the following, which was creating using simulated data:


```